Manuscript

### 2019-01-04 manuscript.Rmd
update manuscript.Rmd for reproducibility in results sections

---
title             : "Manuscript"
shorttitle        : "Subtitle"

@@

```{r g2p-data, include=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
lineage_converter <- unique(ccl_converter[, c("grouped_general_doid", "group_general_lineage_name")])
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"

@@

# print("Drug-Gene-Lineage")
g2p_dgl_tally <- g2p_filt %>% group_by(Drug, Gene, Drug_Gene, group_general_lineage_name) %>% tally() %>% drop_na()
nrow(g2p_dgl_tally)
length(unique(g2p_dgl_tally$Drug))
length(unique(g2p_dgl_tally$Gene))
unique(g2p_filt$Phenotype.Description)
```

@@

@@

After filtering, `r nrow(g2p_filt)` level-A G2P interpretations were carried forward in subsequent analyses, constituting `r length(g2p_drugs)` unique drugs and `r length(g2p_genes)` unique genes in `r length(g2p_druggenes)` distinct level-A combinations (*Figure 1A*). The number of level-A interpretations per drug ranged from `r range(g2p_drug_tally$n)[1]` to `r range(g2p_drug_tally$n)[2]` ($M =$ `r printnum(mean(g2p_drug_tally$n))`, $SD =$ `r printnum(sd(g2p_drug_tally$n))`), `r range(g2p_gene_tally$n)[1]` to `r range(g2p_gene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_gene_tally$n))`, $SD =$ `r printnum(sd(g2p_gene_tally$n))`) per gene, and for each level-A drug-gene combination, the range was `r range(g2p_druggene_tally$n)[1]` to `r range(g2p_druggene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_druggene_tally$n))`, $SD =$ `r printnum(sd(g2p_druggene_tally$n))`). These unique drugs, genes, and combinations are hereafter referred to as G2P drugs, G2P genes, and G2P associations. Of the `r length(g2p_drugs)` unique G2P drugs, 37 of them were screened in at least one of the drug screen datasets, 14 were screened in two of the three datasets, and only three drugs were screened in all three datasets. The remaining 20 un-screened G2P drugs suggests an incompleteness of these drug screens, though these projects frequently release new screening data that may include these drugs in future releases.

```{r}
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
```

@@

In the `r nrow(g2p_filt)` level-A G2P interpretations, there was a wide variation in the specificity of cancer description, ranging from “Waldenström macroglobulinemia” and “hyper eosinophilic advanced syndrome” on the more detailed end to “cancer” on the broad end. Similarly, the drug screen datasets had a wide range in lineage specificity, e.g. CCLE, GTRP, and GDSC all had lineages labeled “leukemia” and “T-cell childhood acute lymphocytic leukemia” and the CRISPR dataset had cell lines labeled “Epstein-Barr virus-related Burkitt lymphoma” and “lymphoma.” To more easily conduct comparative analyses in a cancer-specific manner, we developed a more general lineage grouping method derived from the Human Disease Ontology identity codes (DOIDs, http://disease-ontology.org/) [@schriml2019] assigned to each cell line, yielding 27 unique lineage groupings. This custom lineage grouping was included in our harmonized cancer cell line database (see METHODS).

In the level-A G2P interpretations, there were `r nrow(g2p_dgl_tally)` unique drug-gene-lineage combinations (DGLs), representing seven of our curated cancer lineages: cancer, breast cancer, leukemia, lung cancer, ovarian cancer, skin cancer, and thyroid cancer. CCLE had data for nine DGLs, CTRP had data for 61, and GDSC had data for 53, yielding a total of 73 unique DGLs across all three datasets.

@@

To preclude variation in genomic feature calls across the datasets, we used identical annotations from CCLE for all cell lines screened in the three datasets. The CCLE provided mutation annotations in 19,280 genes across 1,596 cell lines (dataset: depmap_18Q4_mutation_calls.csv, accessed 14 November 2018). CCLE called substitutions using MuTect [@cibulskis2013] and annotated variants using Oncotator [@ramos2015a] and indels using Indelocator (https://software.broadinstitute.org/cancer/cga/indelocator). We filtered mutation calls for point mutations, defined as single-nucleotide insertions, deletions, and substitutions, regardless of result, i.e. frameshift or missense mutation. For this analysis, cell lines harboring non-silent point mutations were considered “mutant” for the gene in question. All other mutations and genes without annotation in the Mutation Annotation Format (MAF) file were considered “wildtype.” “Mutant” and “wildtype” calls were binarily encoded per-gene/per-cell line, regardless of quantity of harbored mutations per gene in a given cell line.

@@

For *Figure 1*, all data were imported and manipulated using pandas (v0.23.0) [@mckinney2011] and plotted using seaborn (v0.90) [@waskom2012], which sat on top of matplotlib (v2.2.2) [@hunter2007] and was executed on Python (v2.7.15) (Python Software Foundation, https://www.python.org/) in an IPython (v5.5.0) [@perez2007] kernel within a localhost Jupyter (v5.2.2) notebook [@kluyver2016]. All other analyses were conducted in R (v3.5.0) (R Foundation for Statistical Computing, https://www.R-project.org/) in an RStudio environment (v1.1.447) [@rstudio2012]. All other plots were drawn using the ggplot2 (v3.0.0) R package [@wickham2016].

@@

Wilcoxon tests were computed using the ggpubr (v.`r print(packageVersion("ggpubr"))`) R package [@kassambara2018].

>>>>

---
title             : "Manuscript Draft"
shorttitle        : "Subtitle"

@@

```{r g2p-data, echo=FALSE, include=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
lineage_converter <- unique(ccl_converter[, c("grouped_general_doid", "group_general_lineage_name")])
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"

@@

# print("Drug-Gene-Lineage")
g2p_dgl_tally <- g2p_filt %>% group_by(Drug, Gene, Drug_Gene, group_general_lineage_name) %>% tally() %>% drop_na()
nrow(g2p_dgl_tally)
length(unique(g2p_dgl_tally$Drug))
length(unique(g2p_dgl_tally$Gene))
g2p_lins <- as.character(unique(g2p_dgl_tally$group_general_lineage_name))
```

```{r drug-data, echo=FALSE, include=FALSE}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele_Length == 1 | is.na(ccle_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)
ccle_data_ptmuts$Dataset <- "CCLE"
ccle_data_ptmuts$Hugo_Symbol <- ordered(ccle_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele_Length == 1 | is.na(ctrp_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)
ctrp_data_ptmuts$Dataset <- "CTRP"
ctrp_data_ptmuts$Hugo_Symbol <- ordered(ctrp_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele_Length == 1 | is.na(gdsc_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)
gdsc_data_ptmuts$Dataset <- "GDSC"
gdsc_data_ptmuts$Hugo_Symbol <- ordered(gdsc_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ccle_data_g2p$Mutation_Status_Nonsilent <- factor(ccle_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ctrp_data_g2p$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_druggenes)
gdsc_data_g2p$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drug-gene-lineage associations
ccle_data_g2p_dgl <- filter(ccle_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ccle_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_dgl <- filter(ctrp_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ctrp_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_dgl <- filter(gdsc_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
gdsc_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drugs and genes
ccle_data_g2p_grid <- filter(ccle_data_ptmuts, Drug %in% g2p_drugs)
ccle_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_grid <- filter(ctrp_data_ptmuts, Drug %in% g2p_drugs)
ctrp_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_grid <- filter(gdsc_data_ptmuts, Drug %in% g2p_drugs)
gdsc_data_g2p_grid$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Dataset counts
drug_summ <- data.frame(Drug = character(), Dataset_Count = integer(), stringsAsFactors = FALSE)
for(drug in g2p_drugs) {
  counter = 0
  if(drug %in% as.character(unique(ccle_data_g2p$Drug))) { counter <- counter + 1 }
  if(drug %in% as.character(unique(ctrp_data_g2p$Drug))) { counter <- counter + 1 }
  if(drug %in% as.character(unique(gdsc_data_g2p$Drug))) { counter <- counter + 1 }
  drug_summ <- rbind(drug_summ, data.frame(Drug = drug, Dataset_Count = counter))
}
```

@@

```{r ccldb, echo=FALSE, include=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
```

@@

After filtering, `r nrow(g2p_filt)` level-A G2P interpretations were carried forward in subsequent analyses, constituting `r length(g2p_drugs)` unique drugs and `r length(g2p_genes)` unique genes in `r length(g2p_druggenes)` distinct level-A combinations (*Figure 1A*). The number of level-A interpretations per drug ranged from `r range(g2p_drug_tally$n)[1]` to `r range(g2p_drug_tally$n)[2]` ($M =$ `r printnum(mean(g2p_drug_tally$n))`, $SD =$ `r printnum(sd(g2p_drug_tally$n))`), `r range(g2p_gene_tally$n)[1]` to `r range(g2p_gene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_gene_tally$n))`, $SD =$ `r printnum(sd(g2p_gene_tally$n))`) per gene, and for each level-A drug-gene combination, the range was `r range(g2p_druggene_tally$n)[1]` to `r range(g2p_druggene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_druggene_tally$n))`, $SD =$ `r printnum(sd(g2p_druggene_tally$n))`). These unique drugs, genes, and combinations are hereafter referred to as G2P drugs, G2P genes, and G2P associations. Filtering our screening datasets by these clinical criteria yields panels of potentially clinically relevant omic features. Of the `r length(g2p_drugs)` unique G2P drugs, `r nrow(drug_summ[drug_summ$Dataset_Count >= 1,])` of them were screened in at least one of the drug screen datasets, `r nrow(drug_summ[drug_summ$Dataset_Count >= 2,])` drugs were screened in at least two of the three datasets, and only `r nrow(drug_summ[drug_summ$Dataset_Count >= 3,])` were screened in all three datasets. The remaining `r nrow(drug_summ[drug_summ$Dataset_Count == 0,])` un-screened G2P drugs suggests an incompleteness of these drug screens, though these projects frequently release new screening data that may include these drugs in future releases.

```{r grid-wilcox-res, echo = FALSE}
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
```

@@

In the `r nrow(g2p_filt)` level-A G2P interpretations, there was a wide variation in the specificity of cancer description, ranging from “Waldenström macroglobulinemia” and “hyper eosinophilic advanced syndrome” on the more detailed end to “cancer” on the broad end. Similarly, the drug screen datasets had a wide range in lineage specificity, e.g. CCLE, GTRP, and GDSC all had lineages labeled “leukemia” and “T-cell childhood acute lymphocytic leukemia” and the CRISPR dataset had cell lines labeled “Epstein-Barr virus-related Burkitt lymphoma” and “lymphoma.” To more easily conduct comparative analyses in a cancer-specific manner, we developed a more general lineage grouping method derived from the Human Disease Ontology identity codes (DOIDs, http://disease-ontology.org/) [@schriml2019] assigned to each cell line, yielding `r length(unique(ccl_converter$group_general_lineage_name))` unique lineage groupings. This custom lineage grouping was included in our harmonized cancer cell line database (see METHODS).

In the level-A G2P interpretations, there were `r nrow(g2p_dgl_tally)` unique drug-gene-lineage combinations (DGLs), representing seven of our curated cancer lineages: `r paste(g2p_lins[1:length(g2p_lins) -1], collapse = ", ")`, and `r g2p_lins[length(g2p_lins)]`. CCLE had data for `r length(unique(ccle_data_g2p_dgl$Drug_Gene_Lin))` DGLs, CTRP had data for `r length(unique(ctrp_data_g2p_dgl$Drug_Gene_Lin))`, and GDSC had data for `r length(unique(gdsc_data_g2p_dgl$Drug_Gene_Lin))`, yielding a total of `r length(union(union(unique(ccle_data_g2p_dgl$Drug_Gene_Lin), unique(ctrp_data_g2p_dgl$Drug_Gene_Lin)), unique(gdsc_data_g2p_dgl$Drug_Gene_Lin)))` unique DGLs across all three datasets. Only `r length(intersect(intersect(unique(ccle_data_g2p_dgl$Drug_Gene_Lin), unique(ctrp_data_g2p_dgl$Drug_Gene_Lin)), unique(gdsc_data_g2p_dgl$Drug_Gene_Lin)))` DGLs were screened in all three datasets,

@@

To preclude variation in genomic feature calls across the datasets, we used identical annotations from CCLE for all cell lines screened in the three datasets. The CCLE provided mutation annotations in 19,280 genes across 1,596 cell lines (dataset: depmap_18Q4_mutation_calls.csv, accessed 14 November 2018). CCLE called substitutions using MuTect [@cibulskis2013] and annotated variants using Oncotator [@ramos2015a] and indels using Indelocator (https://software.broadinstitute.org/cancer/cga/indelocator). We filtered mutation calls for point mutations, defined as single-nucleotide insertions, deletions, and substitutions, regardless of result, i.e. frameshift, missense, or nonsense mutation. For this analysis, cell lines harboring non-silent point mutations were considered “mutant” for the gene in question. All other mutations and genes without annotation in the Mutation Annotation Format (MAF) file were considered “wildtype.” “Mutant” and “wildtype” calls were binarily encoded per-gene/per-cell line, regardless of quantity of harbored mutations per gene in a given cell line.

@@

```{r}
r_v <- paste0("v", getRversion())
# rstudio_v <- paste0("v", RStudio.Version()$version)
ggplot_v <- paste0("v", packageVersion("ggplot2"))
ggpubr_v <- paste0("v", packageVersion("ggpubr"))
```

For *Figure 1*, all data were imported and manipulated using pandas (v0.23.0) [@mckinney2011] and plotted using seaborn (v0.90) [@waskom2012], which sat on top of matplotlib (v2.2.2) [@hunter2007] and was executed on Python (v2.7.15) (Python Software Foundation, https://www.python.org/) in an IPython (v5.5.0) [@perez2007] kernel within a localhost Jupyter (v5.2.2) notebook [@kluyver2016]. All other analyses were conducted in R (`r r_v`) [@rdevelopmentcoreteam2010, https://www.R-project.org/] in an RStudio environment (v1.1.447) [@rstudio2012]. All other plots were drawn using the ggplot2 (`r ggplot_v`) R package [@wickham2016].

@@

Wilcoxon tests were computed using the ggpubr (`r ggpubr_v`) R package [@kassambara2018].
